---
title: 'CA2: Process and steps'
author: "Catarina Lourenço"
date: "27/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CA 2 Instructions
### Programming with R

> You will write a program (R script) which imports, joins and transforms some advertising data. 
Hypothetically this program will run daily on the latest version of the data, so this is not a one-off, discardable analysis script. It should be well documented with comments, and use functions / pipes to make your code as clearly understandable as possible. There are lots of good articles online about how to approach writing good comments. Significant marks are awarded for readability/understandability: see the Rubric for more information.

### Resources

* RStudio development environment 

* Github and Git

* Lecture notes in Brightspace

* [Tidyverse package from CRAN](https://www.tidyverse.org/). In the lab you may need to activate some / all of these individual packages that are part of Tidyverse:

  + library(tibble)
  + library(readr)
  + library(dplyr)
  + library(tidyr)
  + library(magrittr)
  + library(ggplot2)

* [R for data science](http://r4ds.had.co.nz/)



### Dataset
Download a sample data set at this link. This is representative of the data sets that will be analysed daily, although they will be much larger.

***

### Solution 


#### 1. Import
> Extract the sample data-set zip file to a folder. Your program does not need to extract the zip file, but it should import the four files without any modification by any other program.

* Set up the libraries used. 

```{R, echo=FALSE, include=FALSE}
# Install Tidyverse from CRAN
# install.packages("tidyverse")

# Load necessary libraries
library(tibble)
library(readr)
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
#library(ggplot2)

```

* Importing files with correct column specifications:

```{R, echo=TRUE, include=TRUE}
# Load files into tibbles (using readr functions)
# col_types(): Manually define data type of each column as necessary

advertisers <- read_csv("data/advertiser.csv", 
                        col_types = 
                          cols(
                            ID = col_character() #amend parser guess (double) to string
                            )
                        )

campaigns <- read_csv("data/campaigns.csv" , 
                        col_types = 
                          cols(
                            id = col_character(),
                            advertiser_id = col_character()
                            )
                      )

clicks <- read_tsv("data/clicks.tsv", 
                        col_types = 
                          cols(
                            campaign_id = col_character(),
                            time = col_time(format = "%H:%M:%S"),
                            date = col_date(format = "%d/%m/%Y")
                            #date format needs to be specified because the original doesn't conform to ISO standard
                            
                            )
                   )

impressions <- read_tsv("data/impressions.tsv", 
                        col_types = 
                          cols(
                            campaign_id = col_character(),
                            time = col_time(format = "%H:%M:%S"),
                            date = col_date(format = "%d/%m/%Y")
                            )
                        )

```

See readr's (reference on parsing dates and times)[https://readr.tidyverse.org/reference/parse_datetime.html] 

***

#### 2. Transform
> _Impression_ and _click_ data both have date, time and timezone variables. Transform dates and times so that all data that the program outputs is in the UTC time zone.
You must write the logic for this transformation yourself. Using a 3rd party package such as lubridate will be heavily penalised (see Rubric).

> You can use existing R functions to parse the elements out of a timestamp e.g. hour or date, however you must write the code that adds/subtracts hours, days etc. 

Three columns of interest: time, date and timezone.

* 2.1. Filter rows by timezones:

  + 3 types of timezone: UTC, Eastern Time (EST) and Pacific Time (PST). 
  + PST time zone offset is UTC-08. 
  + EST is UTC-05. 
  + To transform these to UTC, add 8 or 5 hours.


* 2.2. Parse dates and times to numbers.

* 2.3. Transform times and dates:

  + Change time: Add time, use modulo to control the amount of hours that go over to the next day; amount of time is a condition of timezone.
  
  + Change date: Use a calendar function to get the next day (might be a different month, a leap year...) IF there are remainder hours (over 24). Base R has as.Date() function.
  
  + Change timezone



```{R, echo=TRUE, include=TRUE}

# get_hour2 <- function(timestamp){
#   return(unlist(strsplit(as.character(timestamp["time"]), "[:]"))[1])
# }

get_hour <- function(x){
  format(strptime(x, "%H:%M:%S"),'%H')
}

get_mins_secs <- function(x){
  format(strptime(x, "%H:%M:%S"),'%M:%S')
}

#new_clicks <- mutate(clicks, hours = format(strptime(clicks$time, "%H:%M:%S"),'%H'))
# 
# new_clicks <- mutate(new_clicks, hours2 = EST_to_UTC(new_clicks$hours))

# EST_to_UTC <- function(x){
#   result <- as.numeric(get_hour(x["time"])) + 5
#   if(result %% 24){
#     result <- (result %% 24)
#   }
#   return(str(result))
# }

# hours <- apply(new_clicks, 1, EST_to_UTC)
# new_clicks$hours2 = hours

time_format <- function(hour, x) {
    if(hour >= 24){ #if the hour is 24 or more, then it restarts from 0 (and date is +1, handled elsewhere)
      hour <- (hour %% 24)
    }
    new_time <- paste(str_pad(hour, width=2, pad="0"), get_mins_secs(x["time"]),  sep=":")
    return(new_time)
}

# x is the row
convert_timezone <- function(x){
  hour <- x["time"]
  if(x["timezone"] != "UTC"){
    hour <- as.numeric(get_hour(hour))
    if(x["timezone"] == "Eastern time"){
      hour <- hour + 5
    } 
    if(x["timezone"] == "Pacific time"){
      hour <- hour + 8
    }
    hour <- time_format(hour, x)
  }
  return(hour)
}

convert_dates <- function(x){
  is_different_date <- get_hour(x["new_time"]) <  get_hour(x["time"])
  return(ifelse(is_different_date, format(as.Date(x["date"]) + 1),  x["date"]))
}

new_times <- apply(clicks, 1, convert_timezone)
clicks2 <- mutate(clicks, new_time=new_times) 
new_dates <- apply(clicks2, 1, convert_dates)
clicks2$new_dates = new_dates


# as.Date(clicks2$time[2], format="%H:%M:%S") >= as.Date(clicks2$time[2], format="%H:%M:%S")
#apply(clicks, 1, get_hour2) # 1 indicates apply over rows

#hours <- format(strptime(clicks$time,"%H:%M:%S"),'%H')

new_clicks <- mutate(clicks, time=clicks2$new_time, date=clicks2$new_dates, timezone="UTC") %>% left_join(campaigns, by = c("campaign_id" = "id")) %>% left_join(advertisers, by = c("advertiser_id" = "ID")) %>% rename(campaign_name = name.x, advertiser_name = name.y)

write_csv(new_clicks, "clicks_processed.csv")

```


```{r}

new_times_i <- apply(impressions, 1, convert_timezone)
impressions2 <- mutate(impressions, new_time=new_times_i) 
new_dates_i <- apply(impressions2, 1, convert_dates)
impressions2$new_dates = new_dates_i

new_impressions <- mutate(clicks, time=clicks2$new_time, date=clicks2$new_dates, timezone="UTC") %>% left_join(campaigns, by = c("campaign_id" = "id")) %>% left_join(advertisers, by = c("advertiser_id" = "ID")) %>% rename(campaign_name = name.x, advertiser_name = name.y)

write_csv(new_impressions, "impressions_processed.csv")

```

```

#### 3. Join
Add all columns from advertisers and campaigns into the _impression_ and _click_ data.



#### 4. Output
After processing the data the program should write two CSV files _impressions_processed.csv_ and _clicks_processed.csv_